---
title: "Homework 3"
output:
  html_document: default
  pdf_document: default
date: 'Deadline: 9 February, 12:00'
subtitle: 'Linguistic Data: Quantitative Analysis and Visualisation'
---

The solutions should be submitted via GitHub.  

## Part 1. A preliminary training  

*Do not use R (RStudio) to solve problems in Part 1. Answers won't be evaluated. *

### Problem 1

Look at the following histogram and answer the questions.

```{r, echo=FALSE}
set.seed(1234)
x <- rnorm(20, mean=3, sd=6)
hist(x, col="tomato", main="")
```

a. What is the proportion of values in the sample that exceed $5$? Explain your answer.
0.25
1+10+4+4++1=20 (всего)
5/20 = 0.25
b. Indicate the interval where the median of this sample can lie. Explain your answer.
Медиана, скорее всего, будет проходить в промежутке между -5 до 0, потому что  примерно  в этом промежутке сосредотачивается половина (10) от всего количества элементов.  
c. How the histogram will change if we add an element $7$ to the sample? Explain your answer.
Если добавить элемент 7, то столбец, лежащий в промежутке от 5 до 10, станет выше на 1 пункт. 

### Problem 2

a. Look at the histograms of two samples. They illustrate the distribution of normalized average reaction time to frequent words (in ms) in two groups of people.

```{r, echo=FALSE, fig.height=3}
set.seed(1234)
sample1 <- rnorm(100, 500, 50)
sample2 <- rnorm(100, 500, 20)
hist(sample1, col="lightblue", xlim=c(100, 900))
```

```{r, echo=FALSE, fig.height=3}
hist(sample2, col="lightgreen", xlim=c(100, 900))
```

Which of the samples has a larger variance? Explain your answer.
Дисперсия больше в гистограмме 1, потому что в гистограмме 2 вся выборка умещается в интервал от 400 до 600, а в гистограмме 1 выходит за пределы 400 и 600.

\newpage

b. Look at the histograms of two samples.

```{r, echo=FALSE, fig.height=8}
set.seed(12)
sample1 <- rnorm(100, 3, 5)
sample2 <- rnorm(100, 10, 0.8)
par(mfrow=c(2,2))
hist(sample1, col="lightblue", xlim=c(-40, 40))
hist(sample2, col="lightgreen", xlim=c(5, 15))
```

Which of the samples has a larger variance? Explain your answer.
Дисперсия больше на графике 1, т.к. там разброс лежит где-то от -10 до 15, а на графике 2 разброс равен 4 (от 8 до 12).

## Part 2

*Do not use R (RStudio) to solve problems in Part 2. Answers for problem 3 will be evaluated. Please paste YES or NO into (empty) code blocks and explain you answer below the block.*

### Problem 3

Below is the histogram of the number of mistakes students made while writing an examination essay in English. Look at the histogram and answer the questions.

```{r, echo=FALSE, fig.height=4}
set.seed(2)
mistakes <- rbeta(1000, 2, 7)*100
hist(mistakes, breaks=60, col= "deepskyblue", ylim=c(0, 40))
```

### 3.1
Is it true that $50$% students made more than $35$ mistakes?
```
Нет, это не так. 
```
Explain your answer below:
По графику видно, что большая часть закрашенной фигуры лежит в интервале от 0 до 35. Следовательно, более 35 ошибок совершило менее половины всех студентов. 

### 3.2
Is it true that most students made no more than $10$ mistakes?
```
Нет, это не так. 
```
Explain your answer below:
По графику видно, что общая плодщадь закрашенной фигуры, лежащей   интервале от 10 до 80, горадздо больше площади фигуры, лежащей в интервале от 0 до 10. Следовательно, меньше 10 ошибок сделало меньшинство студентов, а остальным стоит подучить английский :)

### 3.3
Which of the following values is closer to be the median of `mistakes`: $10$, $20$, $30$, $40$?
```
Медиана располагается ближе к отметке 20, потому что эта отметкапрмерно делит всех студентов поровну. 
```
Explain your answer below:


### Problem 4. Exact binomial test

The null hypothesis is that $p=0$ (i.e. no success is possible). In a dataset, there is only one success out of 1 000 000 observations. Will you reject the null hypothesis? 
```

```
Explain your answer below:


## Part 3

*Use R (RStudio) to solve problems in Part 3. Your answers will be evaluated. Please paste R code into R code blocks and explain you answer below the block, if needed. *

### Problem 5 

Here is a sample of respondents' age:  

$44$, $50$, $42$, $64$, $66$, $42$, $72$, $56$, $72$, $54$, $46$, $48$, $48$, $52$, $50$, $66$, $84$.

### 5.1
Arrange them in a vector and call it `age`. 
```{r}
age <- c(44, 50, 42, 64, 66, 42, 72, 56, 72, 54, 46, 48, 48, 52, 50, 66, 84)
```

### 5.2
Examine the type of `age` variable (numeric, character, etc).
```{r}
class(age)
```

### 5.3
Plot the histogram of the vector `age` with $5$ bins. Change its color to any you want. (Use either R basic or ggplot2 style for plotting.)
```{r}
hist(age, col= "yellow", breaks = 5)
```


### Problem 6

Here is a series of words:  
*pie, bar, bar, pie, pie, bar, bar, chart*.

### 6.1
Arrange elements above in a vector and call it `words` 
```{r}
words <- c("pie", "bar", "bar", "pie", "pie", "bar", "bar", "chart")
```

### 6.2
Calculate the relative frequences of values in `words` measured in percent.
```{r}
table(words)/sum(table(words))*100
```

## Problem 7. Position of verbs in verses

The dataset [“The last words in verses”](https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/poetry_last_in_lines.csv) is based on texts written in the 1820s and 1920s (Corpus of Russian Poetry of the Russian National Corpus). Authors collected only one line per author to keep observations as independent from each other as possible. 

Variables:  
- Decade — decade of creation: 1820s, 1920s.  
- RhymedNwords — the number of words in the rhyming position (usually one word, but there are two words in cases such as _вина бы_ 'I would like to get) wine' (which is rhymed with _жабы_ 'toad', see http://russian-poetry.ru/Poem.php?PoemId=18261)).  
- RhymedNsyl — the number of syllables in the rhyming position.  
- UPoS — part of speech of the last word. 
- LineText — a sampled verse.  
- Author — the author of the text.  

Can we decide that in verses written in 1920s, verbs in the rhyming posision are used differently (more often or less often) than expected for verbs in general? 

Let's assume that the probability for verbs to be used in any position ('in general') is 17% (according to [](http://www.ruscorpora.ru/new/corpora-stat.html) ).

### 7.1 State hypothesis

What is your null hypothesis $H_0$ and what is the alternative hypothesis $H_1$?
```
H0: Вероятность встретить глаголы в рифме равна 17% (то есть такая же, как и в любой позиции). P(verb in verse) = 17
H1: Вероятность встретить глгол в рифме больше или равна 17% (то есть отличается от обычной ситуации)
P(verb in verse) > 17 or P(verb in verse) < 17
```

### 7.2

Read the dataset [“The last words in verses”](https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/poetry_last_in_lines.csv). Filter out the relevant observations from 1920s, calculate the number of verbs observed in the sample, and the sample size. 
```{r}
#install.packages("readr")
#install.packages("dplyr")
library(readr)
library(dplyr)
verses <- read_tsv("https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/poetry_last_in_lines.csv")
dverses_1920s <- filter(verses, Decade == '1920s')
dverbs_1920s <- filter(verbs_1920s, UPoS == 'VERB')

verbs1920 <- nrow(dverbs_1920s)
verbs1920

sampleSize1920 <- nrow(dverses_1920s)
sampleSize

```

### 7.3

Use an exact binomial test to calculate p-value. 
```{r}
binom.test(verbs1920, sampleSize, p=0.17)
```

### 7.4 Interpret results

Give your interpretation of obtained p-value. Answer the initial question: Can we decide that in verses written in 1920s, verbs are used in the rhyming posision more often or less often than expected?
```
p-value = 22%. это больше, чем 5%. следовательно, мы не отвергаем нулевую гипотезу. Веротность встетить глаголы в рифме не отличается от вероятности встретить глагол в любой другой ситуации. 
```

### 7.5 
*(A bonus problem, extra points in evaluation)*.
Repeat 2.3 for verses written in the 1820s.
```{r}
dverses_1820s <- filter(verses, Decade == '1820s')
dverbs_1820s <- filter(dverses_1820s, UPoS == 'VERB')
verbs1820 <- nrow(dverbs_1820s)
verbs1820
sampleSize1820 <- nrow(dverses_1820s)
sampleSize1820
binom.test(verbs1820, sampleSize1820, p=0.17)

```

Write down your general conlusions about data provided for both 1920s and 1820s data.
```
Вероятность встретить глагол в рифме не превышает вероятность встретиь его в любой другой позиции что в 1820-е, что в 1920-е гг. 
```

### Problem 8. One-sample t-test
Using Icelandic data on vowel duration from seminar [Link](https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/icelandic.csv) test the null hypothesis that the population mean of vowel duration in speaker `shg05` equals 73 (ms).
To perform a one-sample t-test, you can use the following example of R code: 
```
t.test(sample, mu = 7725) # mu is a population mean
```

### 8.1
Write down a two-tailed alternative hypothesis.
```
H1:  The population mean of vowel duration in speaker `shg05` does not equal to 73.
```

### 8.2
Perform a one-sample t-test. 
```{r}
icelandic <- read_csv('https://raw.githubusercontent.com/LingData2019/LingData2020/master/data/icelandic.csv')
ice_shg05 <- filter(icelandic, speaker == 'shg05')
ttest <- t.test(ice_shg05$vowel.dur, mu = 73)
ttest

```

### 8.3
Interpret results. 
```
p-value = 5.1%. Это больше, чем 5%. Значит, мы принимаем нулевую гипотезу. 
```

#### Supplementary reading
Use of exact binomial test in lingiistic research:  

* Gries, Stefan Th. "Phonological similarity in multi-word units." Cognitive Linguistics 22.3 (2011): 491-510. [Link](http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.186.7412&rep=rep1&type=pdf)  
Stefan Gries proves that alliteration is observed in multi-word expressions more often than in general.

* Harald Bayen (2008: 51-52) evaluates the probability of observing exactly one occurrence of the word _hare_ in the corpus sample of 1 mln words given its estimated frequency of 8.23 words per million according to the SELEX frequency database.

On measures of central tendency: 

* Levshina 2015, Chapter 3 (p. 48); Gries 2009, Chapter 1.3 (p. 116). 

On t-test: 

* Gries 2009, Chapter 3 (p. 198). 